### 2024-04-29 thesis/chapters/02-analysis.tex
Use bearing A instead of A+B

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Energy}            & $ E(N) = \sum_{i = 1}^{N} a^2(t) $                                                                    \\ \hline \textit{Energy ratio}                & $E_r = E_i \;/\; \sum_{i = 1}^{N}{E_i} $                                                        \\ \hline
\textit{Spectral roll-on} & $ E(f_c) = 0.05 \cdot E(f_s / 2) $ \\ \hline                    
\textit{Spectral roll-off} & $ E(f_c) = 0.85 \cdot E(f_s / 2) $                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \mathrm{corr}(A_{t-1}, A_t)$ \\ \hline 
\textit{Noisiness}                   & $X_{noise} = \mathrm{SNR}(X) = \mu \;/\; \sigma $                             \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\textit{Spectral negentropy}      & $\Delta I_E(f, \Delta f) = \sum_{k = 0}^{N - 1}{\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}} \cdot \ln\left(\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}\right)$                                                \\ \hline
\end{tabular}
\caption{Frequency-domain features}
\label{tab:fd-features}
\end{table}

@@

``Negentropy measures the inclination of a system to increase its level of organization''~\cite{avoci_spectral_2020}. The larger \emph{spectral negentropy} suggests more fault-induced impulses. $E(k, f, \Delta f)$ denotes the squared envelope spectrum (SES) which is an envelope in the Fourier domain: $\mathcal{F}\{ |y(k, f, \Delta f)|^2 \}$.

If a single principal frequency exists, it can be determined by maximum likelihood estimation. Such frequency would explain the signal spectrum the best~\cite{peeters_large_2004}. The frequency spectrum is a discrete set of amplitudes where peaks have to be reliably identified to create representative attributes.

@@

\subsection{Bearing defect frequencies}
%TODO n = number of balls, d - average of outer ring diameter and outer ring, f_s - rotation speed, beta - contact angle - ball bearings close to 0

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{assets/analysis/bearing-parts.jpg}
    \caption{Bearing parts~\cite{lu_what_is_ball_bearing_2023}}
    \label{fig:bearing-parts}
\end{figure}

@@

\subsection{Empirical wavelet features}
\textbf{Empirical Wavelet Transform} (EWT) constructs adaptive bandpass filters with Meyer wavelet. The inner product of the signal with the scaling function $\hat{\phi_n}$ obtains the approximation coefficients and the inner product with the wavelet function results in detail coefficients.

The normalized frequency axis in range $\omega \in [0, \pi]$ is divided by split points $\omega_0, \dots, \omega_N$ where $\omega_n = f_n \cdot 2 \pi\;/\;f_s$ (Fig.~\ref{fig:ewt-spectrum-segmentation}). Each segment is bounded between $[\omega_{n-1}, \omega_n]$ with a transition phase of width $2\tau_n$ and polynomial transition function $\beta(x)$. A tight frame set of empirical wavelets is built by setting the transition phase proportional to the band boundary: $\tau_n = \lambda \omega_n$, and proportional constant must obey constrain: $\lambda = \min\left(\frac{\omega_{n+1} - \omega_n}{\omega_{n+1} + \omega_n}\right)$. The $N$ boundaries defining different portions of the Fourier spectrum are placed at the center between two consecutive local maxima~\cite{gilles_empirical_2013}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\textwidth]{assets/analysis/EWT.png}
    \caption{Empirical wavelet segmentation of Fourier spectrum~\cite{gilles_empirical_2013}}
    \label{fig:ewt-spectrum-segmentation}
\end{figure}

The drawback of EWT is improper segmentation in noisy and non-stationary signals, producing too many uninformative partitions of the spectrum. The solutions we discuss fall into two groups that apply improvements before or after the wavelet transformation step.

A combination of \textbf{Maximum Correlated Kurtosis Deconvolution} and improved EWT (MKCD-EWT) favors periodic impacts by dynamically constructing an optimal FIR filter that maximizes the correlation kurtosis of the signal. The envelope curve smooths the amplitude spectrum by linear interpolation (Fig.~\ref{fig:mkcd-ewt-segmentation}). The threshold accounting for desired SNR modifies the envelope when more than $n$ components are discovered ($A_h$ is maximal amplitude and $A_l$ is minimal amplitude): $\lambda = A_l + \frac{C}{\mathrm{SNR}}(A_h - A_l)$. \emph{Squared envelope spectrum} and \emph{Teager energy operator spectrum} are computed as fault features for the IMF with the highest kurtosis~\cite{li_fault_2019}.

The segmentation boundary detection can also be accomplished on the spectral envelope consisting of \textbf{Piecewise Cubic Hermite Interpolating Polynomial} (PCHIP-EWT) instead of scanning the Fourier spectrum directly (Fig.~\ref{fig:pchip-ewt-segmentation}). Divisions of the upper cut-off frequency are kept if local power indicates a subband with useful information ($p_i(f) \geq \lambda$)  as in the case of an abrupt level change. Local power is the ratio of $n$th local envelope maximum and the difference between indexes of its adjacent local minima:  $p_i(f) = K_{max(i)(f)} \;/\;\left(f_{k_{min(i+1)}} - f_{k_{min(i)}}\right)$. The best $\lambda$ is set experimentally~\cite{zhuang_improved_2020}.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/MKCD-EWT.png}
        \caption{MKCD-EWT~\cite{li_fault_2019}}
        \label{fig:mkcd-ewt-segmentation}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/PCHIP-EWT.png}
        \caption{PCHIP-EWT~\cite{zhuang_improved_2020}}
        \label{fig:pchip-ewt-segmentation}
    \end{subfigure}
    \caption{Illustration of improved EWT spectral segmentations}
\end{figure}

The inspiration for filtering on an adaptive basis comes from \textbf{Empirical mode decomposition} (EMD). This procedure locates local maxima and minima in the time waveform that are interpolated with cubic splines to form signal envelopes. The average of the upper and lower envelope is repeatedly subtracted from the residuals recovering the higher order and lower frequency IMFs~\cite{wang_computational_2014}.

The EMD is not recommended for practical applications because it suffers from mode mixing problems and lacks rigid theoretical foundations. Many improvements based on the mode sifting process, such as Local Mean Decomposition (LMD), Ensemble Empirical Mode Decomposition (EEMD), Concealed Component Decomposition (CCD)~\cite{tiwari_novel_2021}, and Empirical Wavelet Transform, have been devised as being capable of estimating more reasonable modes.

@@

\textbf{Half-Space Tree} (HS-Tree) stands upon the concept of Isolation forest. It assumes that random splitting of each axis in the feature space will isolate outliers to their separate divisions sooner than non-deviant observations~\cite{gervasi_anomaly_2020, torres_automatic_2022}. This ensemble of trees is better suited for batch setting. HS-Tree stands out in adapting to changing streams because it is trained solely on normal data, requires constant memory, and is faster than density-based methods~\cite{tan_fast_2011}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\textwidth]{assets/analysis/HS-Tree.png}
    \caption{Half-space tree~\cite{tan_fast_2011}}
\end{figure}

A full binary tree is built before the novelty detection begins by splitting tree nodes along the divisions in the randomly chosen perpendicular planes. The node stores its depth, value limits of the axis bisection (half-space), count of contained data points (mass) in two consecutive windows, and links to both child nodes~\cite{tan_fast_2011}.

>>>>

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Energy}            & $ E(N) = \sum_{i = 1}^{N} a^2(t) $                                                                    \\ \hline \textit{Energy ratio}                & $E_r = E_i \;/\; \sum_{i = 1}^{N}{E_i} $                                                        \\ \hline
\textit{Spectral roll-on} & $ E(f_c) = 0.05 \cdot E(f_s / 2) $ \\ \hline                    
\textit{Spectral roll-off} & $ E(f_c) = 0.85 \cdot E(f_s / 2) $                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \mathrm{corr}(A_{t-1}, A_t)$ \\ \hline 
\textit{Noisiness}                   & $X_{noise} = \mathrm{SNR}(X) = \mu \;/\; \sigma $                             \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\textit{Spectral negentropy}      & $\Delta I_E(f, \Delta f) = \sum_{k = 0}^{N - 1}{p \cdot \ln(p); p = \frac{\mathrm{SES}^2}{\frac{1}{N} \sum_{k=0}^{N-1} \mathrm{SES}^2}}$                                                \\ \hline
\end{tabular}
\caption{Frequency-domain features}
\label{tab:fd-features}
\end{table}

@@

``Negentropy measures the inclination of a system to increase its level of organization''~\cite{avoci_spectral_2020}. The larger \emph{spectral negentropy} suggests more fault-induced impulses. The squared envelope spectrum (SES) is interpreted as $E(k, f, \Delta f)$ which is an envelope in the Fourier domain: $\mathcal{F}\{ |y(k, f, \Delta f)|^2 \}$.

\subsection{Features in harmonic frequencies}
If a single principal frequency exists, it can be determined by maximum likelihood estimation. Such frequency would explain the signal spectrum the best~\cite{peeters_large_2004}. The frequency spectrum is a discrete set of amplitudes where peaks have to be reliably identified to create representative attributes.

@@

\subsection{Bearing defect frequencies}

%TODO n = number of balls, d - average of outer ring diameter and outer ring, f_s - rotation speed, beta - contact angle - ball bearings close to 0

% Rolling element bearings have four very important components, two circular races, between which the rolling elements—rollers, tapered rollers, or spherical balls—are held. The fourth important element in a rolling element bearing is the cage or the separator. The cage ensures that no two rolling elements come in contact with each other
% The bearing vibration is further complicated when the radial loads on the bearing change during rotation of the shaft, and defects like impurities or scratches, pits, and so on, develop on the bearing element surfaces
% The characteristic bearing defect frequencies for the different rolling elements of bearings can be determined

\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{assets/analysis/bearing-parts.jpg}
    \caption{Bearing parts~\cite{lu_what_is_ball_bearing_2023}}
    \label{fig:bearing-parts}
\end{figure}

@@

@@

%\cite{tan_fast_2011}.

### 2024-04-29 thesis/chapters/03-design.tex
Use bearing A instead of A+B

\section{Research questions}
% Questions
% Goals
% Project risk management
This thesis aims to provide answers to four research questions. The focus is primarily on making data flow more efficient in an industrial sensor network that monitors rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which temporal and spectral features can be extracted from vibration signals to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput when chosen signal features are used?
	\item What accuracy of prediction models can be achieved with various feature subsets?
	\item How can machinery faults be continuously identified and predicted based upon collected events?
\end{enumerate}

@@

However, we leave out from our efforts experiments on the data features calculated from wavelets and peaks in the spectral domain. The reason is that we did not find a way to represent extracted features more succinctly as a single number. We also did not discover a strategy for choosing only the relevant frequency bins. The assembled description is retained to lead further research on that topic.

The goals are impacted by certain \textbf{risks} that are assessed and tracked:

\begin{enumerate}[label=R-\arabic*., font=\bfseries]
\itemsep0pt
\item Machines in the real environment will not be available for vibration measurements. \\ \emph{Mitigation}: Contact and establish collaboration with alternative partners. 
\item Repeated measurements will not be consistent, and a lack of data is obtained from different classes.
\\ \emph{Mitigation}: Prepare a plan for measurements and photo document sensor placements. 
\item Fault modes can not be reliably differentiated and labeled in the dataset.
\\ \emph{Mitigation}: Consult domain experts and machine maintenance staff.
\item Suggestions made by exploring the MaFaulDa dataset will not be applicable in practice. \\
\emph{Mitigation}: When applicable, apply the same procedures as in MaFaulDa to measured data.
\end{enumerate}

@@

\section{Exploratory data analysis of MaFaulDa}
% Fault counts
\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
\hline
\textbf{Bearings} & \textbf{Severity} & \textbf{Misalignment} & \textbf{Imbalance} & \textbf{Cage fault} & \textbf{Ball fault} & \textbf{Outer race fault} & \textbf{Normal} & \textbf{$\Sigma$} \\ \hline
A                 & No                & 498                   & 333                & 188                 & 186                 & 184                       & 49              & 1438              \\ \hline
B                 & No                & 498                   & 333                & 188                 & 137                 & 188                       & 49              & 1393              \\ \hline
A+B               & No                & 996                   & 666                & 376                 & 323                 & 372                       & 98              & 2831              \\ \hline
A                 & Yes               & 248                   & 188                & 91                  & 87                  & 86                        & 738             & 1438              \\ \hline
B                 & Yes               & 248                   & 188                & 90                  & 45                  & 90                        & 732             & 1393              \\ \hline
A+B               & Yes               & 496                   & 376                & 181                 & 132                 & 176                       & 1470            & 2831              \\ \hline
\end{tabular}
\end{table}

@@

% Features Range
% Mafaulda
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-TD-dim1-A+B.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-FD-dim1-A+B.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-TD-dim3-A+B.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-FD-dim3-A+B.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{Feature range in MaFaulDa}
\end{figure}

@@

% HW parameters and mention temporary device
\begin{table}[ht]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & STMicroelectronics \\ \hline
\textbf{Bus}                                     & SPI                \\ \hline
\textbf{Axis}                                    & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 2 - 16      \\ \hline
\textbf{Bandwidth} (kHz)                          & 5 - 6.3            \\ \hline
\textbf{Sensitivity} (mg/LSB)                    & 0.06 - 0.49       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 75                 \\ \hline
\textbf{Microcontroller}                         & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 16                 \\ \hline
\textbf{FIFO}                                    & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Hardware parameters of previous and proposed sensor units}
\label{tab:design:hw-sensors}
\end{table}

>>>>

\section{Research questions}
This thesis aims to provide answers to four research questions. The focus is primarily on making data flow more efficient in an industrial sensor network that monitors rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which temporal and spectral features can be extracted from vibration signals to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput when chosen signal features are used?
	\item What accuracy of prediction models can be achieved with various feature subsets?
	\item How can machinery faults be continuously identified and predicted based upon collected events?
\end{enumerate}

@@

However, we leave out from our efforts experiments on the data features calculated from wavelets and peaks in the spectral domain. The reason is that we did not find a way to represent extracted features more succinctly as a single number. We also did not discover a strategy for choosing only the relevant frequency bins.

@@

\section{Exploratory data analysis of MaFaulDa}
\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|lr|r|r|r|r|r|r|}
\hline
\multicolumn{2}{|l|}{\textbf{Bearing}}                                            & \multicolumn{1}{l|}{A}  & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}  & \multicolumn{1}{l|}{A}   & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}   \\ \hline
\multicolumn{2}{|l|}{\textbf{Relabel by severity}}                                & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{No}  & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} \\ \hline
\multicolumn{1}{|l|}{\multirow{6}{*}{\textbf{Fault}}} & \textbf{misalignment}     & 498                     & 996                      & 498                     & 248                      & 496                      & 248                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{imbalance}        & 333                     & 666                      & 333                     & 188                      & 376                      & 188                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{cage fault}       & 188                     & 376                      & 188                     & 91                       & 181                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{ball fault}       & 186                     & 323                      & 137                     & 87                       & 132                      & 45                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{outer race fault} & 184                     & 372                      & 188                     & 86                       & 176                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{normal}           & 49                      & 98                       & 49                      & 738                      & 1470                     & 732                      \\ \hline
\multicolumn{2}{|l|}{$\Sigma$}                                                    & \textbf{1438}           & \textbf{2831}            & \textbf{1393}           & \textbf{1438}            & \textbf{2831}            & \textbf{1393}            \\ \hline
\end{tabular}
\caption{Number of rows in MaFaulDa according to label}
\end{table}

@@

% Feature correlations
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/corr-A-3-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/corr-A-3-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{Feature correlations in MaFaulDa}
\end{figure}

% Features Range
% Mafaulda
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-TD-dim1-A.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-FD-dim1-A.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-TD-dim3-A.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/features-FD-dim3-A.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{Feature range in MaFaulDa}
\end{figure}

@@

% HW parameters and mention temporary device
\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & STMicroelectronics \\ \hline
\textbf{Bus}                                     & SPI                \\ \hline
\textbf{Axis}                                    & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 2 - 16      \\ \hline
\textbf{Bandwidth} (kHz)                          & 5 - 6.3            \\ \hline
\textbf{Sensitivity} (mg/LSB)                    & 0.06 - 0.49       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 75                 \\ \hline
\textbf{Microcontroller}                         & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 16                 \\ \hline
\textbf{FIFO}                                    & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Hardware parameters of accelerometer data logger}
\label{tab:design:hw-sensors}
\end{table}
